### 2018-07-09 doc/manuscript.tex
fix fixation controller

\documentclass[11pt,a4paper,draft=false]{report}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{bm,txfonts}       % bm: bold for Greek letters, txfonts -> varmathbb
\usepackage{breqn}              % automatically brake line in equations
\usepackage{cancel}
\usepackage{arydshln}         % matrix dash lines
\usepackage{booktabs}         % tables
\usepackage{color,soul}         % used for \hl{}
\usepackage[inline]{enumitem}   % use to set labels in the enumerate environment
\usepackage{fullpage}
\usepackage[pdftex]{graphicx}
\graphicspath{{./figures/}{./tikz/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{caption}
\captionsetup{labelfont=bf}
\usepackage[caption=false,labelfont=bf]{subfig}
\usepackage{listings}
% \usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{url}
\usepackage[pdfpagelabels,pdfusetitle,colorlinks=true,pdfborder={0 0
  0},pdfauthor={Dimitar Stanev}]{hyperref}
\usepackage{natbib}
\bibliographystyle{IEEEtranN}

% indexing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \usepackage{makeidx}
% \makeindex

% make use of \index{word} in text

% acronyms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[acronym]{glossaries}

\newacronym{dofs}{DoFs}{Degrees of Freedom}
\newacronym{dof}{DoF}{Degree of Freedom}
\newacronym{eoms}{EoMs}{Equations of Motion}
% \newacronym{so}{SO}{Static Optimization}
\newacronym{fd}{FD}{Forward Dynamics}
\newacronym{pd}{PD}{Proportional Derivative}
\newacronym{cns}{CNS}{Central Nervous System}
\newacronym{lr}{LR}{Lateral Rectus}
\newacronym{sr}{SR}{Superior Rectus}
\newacronym{ir}{IR}{Inferior Rectus}
\newacronym{mr}{MR}{Medial Rectus}
\newacronym{so}{SO}{Superior Oblique}
\newacronym{io}{IO}{Inferior Oblique}
\newacronym{rmse}{RMSE}{Root Mean Square Error}
\newacronym{em}{EOMs}{Extraocular Muscles}
\newacronym{fl}{F-L}{Force-Length}
\newacronym{fv}{F-V}{Force-Velocity}
\makeglossaries{}

% custom commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n} % mathcal bf
\newcommand{\ten}[1]{\mathbfcal{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\renewcommand*{\vec}[1]{\bm{#1}}
\newcommand{\R}[1]{\mathfrak{R}^{#1}}
\newcommand{\inr}[1]{\in\R{#1}}

\newcommand{\subfigureautorefname}{Figure} % subfloat
\renewcommand*{\figureautorefname}{Figure}
\renewcommand*{\sectionautorefname}{Section}
\renewcommand*{\subsectionautorefname}{Subsection}
\newcommand{\algorithmautorefname}{Algorithm}
\makeatletter
% \renewcommand\thealgorithm{\thechapter.\arabic{algorithm}}
\@addtoreset{algorithm}{chapter}
\makeatother
\def\equationautorefname~#1\null{Equation #1\null}

% in order to add newline in table cells
\usepackage{makecell}
\renewcommand\theadalign{bc}
\renewcommand\theadfont{\bfseries}
\renewcommand\theadgape{\Gape[4pt]}
\renewcommand\cellgape{\Gape[4pt]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\title{An Open-Source OpenSim Oculomotor Model for Kinematic and Dynamic
  Simulation}

\author{Constantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
    Engineering Department, University of Patras, Greece, Corresponding author:
    stanev@ece.upatras.gr} and Konstantinos Moustakas}

\date{\today}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  bla
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section*{Introduction}\label{sec:introduction}

Rapid and accurate eye movements are crucial for coordinated direction of
gaze. Studying human eye movement has significant implications for improving our
understanding of the oculomotor system and treating vasomotor visuomotor disorders. Over
the past decades, biomechanics simulation has provided the means to analyze
different human movements. The same principles can be used to analyze visual
tasks by modeling the musculoskeletal properties of the oculomotor
system. Consequently, this model can be used to investigate muscle activation
patterns during static fixation, analyze the dynamics of various eye movements,
calculate metabolic costs and simulate different eye disorders, such as
different forms of strabismus. Last but not least, it can be easily integrated
with available full body models in order to analyze the relation between the
vestibular and oculomotor systems.

Eye movements are a result of the coordination of the six \gls{em}. Clinical
trials have provided a profound knowledge of how the \gls{em} act on the eye
globe, the resistive tension of the surrounding tissues and the length-tension
relationship of the muscles. Various computational models of the extraocular
muscles and orbital mechanics have been proposed, which provide insight and
scientific bases for oculomotor biomechanics, control of eye movement and
binocular misalignment. These models focus on the realism of muscle behavior and
they were based on the viscoelastic properties and physiological data \gls{em}.

The first 3D biomechanical model was developed by~\cite{Robinson1964a,
  Robinson1969}, who simplified the formulation by considering the elasticity of
the \gls{em} ignoring their dynamics. The model incorporates anatomically
realistic muscle paths and empirical innervation-length-tension
relationships. To study the neural control of rapid saccadic movement, models
using anatomical and mechanical properties of \gls{em} have been developed. They
take account that the actual force is a complex nonlinear function of the muscle
length, velocity, and innervation~\cite{Thelen2003, Millard2013}. Such models
have the advantage of supporting 3D dynamic simulations and have been used to
analyze neural controllers and the pulley hypothesis.

We introduce a detailed open-source 3D biomechanical model of the human eye,
which can be used for dynamic simulation\footnote{SimTK project:
  \url{https://simtk.org/projects/eye}}, based on the data reported
in~\cite{Iskander2018}. The model is implemented in
\texttt{OpenSim}~\cite{Delp2007}, which is an open-source framework for modeling
and simulation of musculoskeletal systems. The architecture and dynamic muscle
properties are based on physiological and kinematic measurements of the human
eye. The model incorporates an eye-globe, orbital suspension tissues and six
\gls{em}. Using this model we are able to assess the excitation and activation
patterns for a variety of targets by applying a closed-loop fixation controller
that drives the model to perform saccadic movements in a \gls{fd} manner. The
controller minimizes the error between the desired saccadic trajectory
(reference input) and the predicted movement (calculated output).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section*{Methods}\label{sec:methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection*{Eye Modeling}\label{sec:eye-Modeling}

The orbital plant consists of the globe (eyeball), three pairs of extraocular
muscles, and connective tissues. The size of an emmetropic human adult eye is
approximately 0.0242 m (transverse, horizontal), 0.0237 mm (sagittal, vertical),
0.022–0.0248 m (axial, anteroposterior) with no significant difference between
sexes and age groups. In the transverse diameter, the eyeball size may vary from
0.021 m to 0.027 m. Thus, it can be approximated by a solid sphere with 0.012 m
radius. The eyeball was constructed in \texttt{Blender}, an open-source software
3D creation software. We used a spherical mesh with 32 segments and 12 rings, to
construct the vitreous humor (body) as solid sphere and a conical plate to
construct the cornea. The weight of an average human eye is 0.0075 kg and the
moment of inertia can be calculated similarly as in the case of a spherical
homogeneous and isotropic object with radius 0.012 m ($\mat{I} = 2/5 m r^2$ at
the center of mass).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection*{Muscle Modeling}\label{sec:muscle-modeling}

The six \gls{em}, including four rectus muscles and two oblique muscles, are
controlled by the cranial nerves so as to track a visual target and to stabilize
the image of the object. The \gls{lr} and \gls{mr} muscles form an antagonistic
pair to produce horizontal eye movements. The \gls{sr} and \gls{ir} muscles form
the vertical antagonist pair, which mainly controls vertical eye movement and
also affects rotation about the horizontal plane and the line of sight
(secondary action) due to insertion positions and the path of the muscles. The
\gls{so} muscle passes through the cartilaginous trochlea attached to the
orbital wall, which reflects the \gls{so} path by 51 deg. The \gls{io} muscle
originates from the orbital wall anteroinferior to the globe center and inserts
on the sclera posterior to the globe equator. The primary actions of \gls{so}
and \gls{io} cause rotation of the globe around the visual axis and vertical
movement.

The model relies on the passive pulley assumption in order to keep it simple and
provide faster simulation speed. \autoref{tab:muscle-path} shows the positions
of muscle pulleys, as well as the origin and insertion points of the \gls{em},
defined in the local body coordinates of the globe. The data is based on
physiological measurements~\cite{Iskander2018}, with some minor adaptation so as
to prevent penetration into the eye globe. Since no position was documented for
the origin of the \gls{so}, a point close to the origins of the rectus muscles
was chosen to match the fiber length in the primary position of the \gls{so}
muscle.

\begin{table}[h]
  \caption{Muscle path for the six \gls{em} (dimensions are given in
    meters).}\label{tab:muscle-path}
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \textbf{Muscle}
    & \multicolumn{3}{c}{\textbf{Origin}}
    & \multicolumn{3}{c}{\textbf{Pulley}}
    & \multicolumn{3}{c}{\textbf{Insertion}} \\
    \midrule
    & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}}
    & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}}
    & \textit{\textbf{Ix}} & \textit{\textbf{Iy}} & \textit{\textbf{Iz}} \\
    \midrule
    \gls{lr} & -0.034 & 0.0006 & -0.013 & -0.0102 & 0.0003 & 0.012 & 0.0065 & 0 & 0.0101 \\
    \gls{mr} & -0.030 & 0.0006 & -0.017 & -0.0053 & 0.00014 & -0.0146 & 0.0088 & 0 & -0.0096 \\
    \gls{sr} & -0.0317 & 0.0036 & -0.016 & -0.0092 & 0.012 & -0.002 & 0.0076 & 0.0104 & 0 \\
    \gls{ir} & -0.0317 & -0.0024 & -0.016 & -0.0042 & -0.0128 & -0.0042 & 0.00805 & -0.0102 & 0 \\
    \gls{so} & 0.0082 & 0.0122 & -0.0152 & -0.030834 & 0.001145 & -0.01644 & 0.0044 & 0.011 & 0.0029 \\
    \gls{io} & 0.0113 & -0.0154 & -0.0111 & -0.00718 & -0.0135 & 0 & -0.008 & 0 & 0.009 \\
    \bottomrule
  \end{tabular}
\end{table}

The Millard muscle model~\cite{Millard2013} has been adopted for the modeling of
the \gls{em}, which allows to manually fit the reported data for the \gls{fl}
curves. The muscles were modeled using the rigid tendon assumption that ignores
the elasticity of the tendon. This means that the series element of the muscle
model is not included (the tendon length $l^T$ is equal to the tendon slack
length $l_s^T$). This assumption is valid when the ratio of the tendon length to
the muscle length is less or equal to one, as the in the case of all
\gls{em}. \gls{em} are considered parallel-fibered muscles, so the pennation
angle is zero ($\alpha = 0$). Maximum isometric force $f_o^M$, optimal fiber
length $l_o^M$ and tendon length $l^T$ are presented
in~\autoref{tab:muscle-path}.

The active and passive \gls{fl} curves for the \gls{em} differ from that of a
skeletal muscle. As shown in \autoref{fig:millard-curves}, we can fine-tune
these curves so as to fit the experimental data available for the gls{lr}
muscle. The following values where used for the active \gls{fl} curve:

\begin{itemize}
\item min norm active fiber length: 0.55
\item transition norm fiver length: 0.7
\item max norm active fiver length: 1.8
\item shallow ascending slope: 2.4
\item minimum value: 0.0
\end{itemize}
%
and for the passive \gls{fl} curve accordingly:

\begin{itemize}
\item strain at zero force: -0.18
\item strain at one norm force: 0.4
\end{itemize}

\begin{figure}[!ht]
  \subfloat[Active \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
    keepaspectratio]{active-force-length-curve.png}\label{fig:active-force-length-curve}}
  \subfloat[Passive \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
    keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
  \caption{The active and passive \gls{fl} curve definition of the Millard
    muscle model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
\end{figure}

The parameters that describe the above relationships were chosen to fit the
curves reported in~\cite{Iskander2018}, by matching the \gls{fl} relationship at
maximum activation of the \gls{lr} muscle. This represents the first part of
testing the fidelity of the model. Due to lack of data describing the other
muscles’ \gls{fl} relationships, we used the parameters found describing the
normalized curves of active and passive \gls{fl} relationships of the gls{lr}
for the other \gls{em} as well.

\gls{em} have a higher fraction of fast twitch fibers and thus different
\gls{fv} behavior, due to different structures compared to skeletal
muscles. Despite that, the default Millard \gls{fv} curve was used for the six
\gls{em}, since the behavior of the selected muscle model depends mainly on the
maximum contraction velocity $v^{\text{max}}$. The maximum muscle contraction
velocity is tuned so as to match the peak velocity of saccadic eye movement 15.7
rad / s. Following this definition, the maximum muscle contraction velocity is
given in optimal fiber length per seconds and it is thus different for each
\gls{em}, as their optimal fiber length is different. Furthermore, because of
the different structure of neural control of the eyes, activation and
deactivation delays ($\tau_d = 5$ ms) are lower than in skeletal
muscles. Finally, two separate wrapping spheres for the rectus muscles and the
oblique muscles were created, to avoid abnormal changes on the \gls{fl} curve as
the eyeball rotates in the three directions.

\begin{table}[h]
  \caption{Millard muscle parameters for the
    \gls{em}.}\label{tab:muscle-parameters}
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \thead{Muscle}
    & \thead{Maximum Isometric \\ Force (N)}
    & \thead{Optimal Fiber \\ Length (m)}
    & \thead{Tendon Slack \\ Length (m)}
    & \thead{Maximum Contraction \\ Velocity (m / s)} \\
    \midrule
    \gls{lr} & 1.4710 & 0.04898 & 0.0084 & 3.8483 \\
    \gls{mr} & 1.5740 & 0.04084 & 0.0038 & 4.6155 \\
    \gls{sr} & 1.1768 & 0.04487 & 0.0054 & 4.2009 \\
    \gls{ir} & 1.4269 & 0.04549 & 0.0048 & 4.1437 \\
    \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 3.5863 \\
    \bottomrule
  \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection*{Passive Connective Tissues}\label{sec:passive-connective-tissues}

The passive connective tissues of the eyeball apply a restoring force, which
brings the globe back to the central position when the net force from the
\gls{em} is zero. These tissues include all non-muscular suspensory tissues,
such as Tenon’s capsule, the optic nerve, the fat pad and the conjunctiva. The
force-displacement curve of the net elasticity can be represented as

\begin{equation}\label{equ:passive-tissue}
  \vec{f}_t = -k_p \vec{q} - k_c 10^{-3} \vec{q}^3 - k_d * \vec{\dot{q}}
\end{equation}
% 
where, $\vec{f}_t$ represents the passive tissue forces, $k_p= 0.002225$ N m /
rad, $k_c= 34.5297$ N m / (rad^3) and $k_v= 0.002$ N m s / rad the constants and
$\vec{\dot{q}} \inr{3}$ the rotational coordinates of the model. These forces
serve the eye’s stabilization, and are modeled using \texttt{OpenSim}'s
expression based coordinate force.
% 0.33 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) = 0.18542028707494282 * 0.012
% 1.56 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) ** 3 = 2877.4856893811248 * 0.012

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section*{Results}\label{sec:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection*{Model Validation}\label{sec:model-validation}

To meet good fidelity criteria, a model requires to be verified and
validated. In our study, verification in oculomotor models was performed by
comparing the \gls{fl} characteristic curves of the modeled \gls{em} to
published data. These data include only the lateral rectus characteristic
curves. However, we can make a safe assumption that the other muscles have
similar properties. More characteristic curves showing the \gls{fl} relationship
of the other \gls{em}, as well as the changes in muscle length with rotation in
all directions, are presented in the end of this report. Further verification
can be done by comparing the model’s joint forces produced during horizontal
movement with clinically collected data, but these data are not available.
Validation was performed by simulating and analyzing synthesized eye movements
while ensuring that Listing Law was obeyed by maintaining zero torsion in
secondary gaze positions, and by testing if the assessed innervations are
sufficient in fixating the eye at a desired position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection*{Fixation Controller}\label{sec:fixation-controller}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% \section*{Discussion}\label{sec:discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section*{Conclusion}\label{sec:conclusion}

A realistic ocular model that represents the ocular motility of a normal human
eye was presented. The model can be used to simulate different saccadic
movements and obtain the muscle activations required to track the movement. The
model was validated against experimental measured data and is able to synthesize
saccadic movements reasonably well.

We even showed that the simulation results produced by static optimization and
forward dynamics in OpenSim are less satisfactory than the proposed solution
where we used a \gls{pd} controller to minimize the tracking error between the
desired and estimated trajectory of saccadic movements. The desired kinematics
were tracked within a maximum \gls{rmse} of and in horizontal and vertical
saccades respectively. With the application of the controller, a very rapid
instantaneous acceleration and a constant velocity to sustain clear vision is
attained. The produced activation levels were in accordance with the
descriptions found in the bibliography and the highest activation levels were
shown only during the time intervals when the main agonist muscle was activated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{mylibrary}
\addcontentsline{toc}{chapter}{Bibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: